---
title: "Exploratory Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)

# source all R files
source('summary_information.r')
source('chart2.R')
```

## Project title

Give your project a concise, interesting title that summarizes the entirety of your project

### Authors

Names of your team members

### Date

Winter 2022

### Abstract

No more than three sentences that summarize your project. Focus on the very most important aspects. For example: (1) "Our main question is .... This question is important because .... To address the question, we will ...." (2) "We are concerned with ..., because .... To address this concern, we plan to ...." (3) "Consider that .... This is important because .... Accordingly, we plan to ...."

### Keywords

3-5 keywords that summarize your project.
(e.g., "Keywords: human physiology; bicycle exercise; Gen Z; times-series data")

### Introduction  

Briefly introduce your project.  Include 3-5 research questions. What motivates the questions? Why are they important? (at least 200 words)

### Related Work  

Describe your topic and related work in this space. You must include 3 citations to related work (URLs to similar work, high quality articles from the popular press, research papers, etc. ) Please use a standard citation style of your choice. (at least 200 words)

https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7588920/

### The Dataset

Where did you find the data? Please include a link to the data source  
Who collected the data?  
How was the data collected or generated?  
Why was the data collected?  
How many observations (rows) are in your data?  
How many features (columns) are in the data?  
What, if any, ethical questions or questions of power do you need to consider when working with this data?  
What are possible limitations or problems with this data?   (at least 200 words)

### Implications

Assuming you answer your research questions, briefly describe the expected or possible implications for technologists, designers, and policymakers. (at least 150 words)

### Limitations & Challenges
What challenges or limitations might you need to address with your project idea more broadly? Briefly discuss. (at least 150 words)

### Summary Information

Write a summary paragraph of findings that includes the 5 values calculated from your summary information R script

### Table

Include a table of aggregate information

Describe why you included the table and what information it reveals

### Chart 1

Include a chart

Describe why you chose this chart and what information it reveals

Here's an example of how to run an R script inside an RMarkdown file:



### Chart 2

```{r chart2,warning=FALSE}
nutrition.plot
```

Household income and nutritional intake are closely related. Different foods have different demand and price elasticity. We suspect that when household income is low, people pursue basic energy intake. When people have higher incomes, they make different choices about nutrition to balance their diet. From the plot, we see that the x-axis is family annul income, but to better explore the linear relation between the income and three common nutrition, giving each income level a numeric value is better to see the trend. Below is sample of the meaning of numeric represents of family annual income with increment of $4,999. 

|category|meaning|
|:------:|:-----:|
|1| \$0 to $4,999|
|2| \$5,000 to $9,999|

And, we choose three most common nutrition: Carbohydrate, protein, and fat measured in gram and plotted as y-axis. We see that as income increases, all of them decrease, but the negative trend between carb and income is the most significant. We find that carbohydrate can be viewed as necessity at low income levels, people at which cannot afford better and diverse food choices. However, at higher income level, a higher negative elasticity appears between income and carbohydrate(*(Yongjian, 2020)*)[https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7588920/]. They have more choices and they may choose different nutrition intake to balance their food rather than just solve the probelm of hunger. 

### Chart 3

Include a chart

Describe why you chose this chart and what information it reveals
